---
title: "Week 12 Exercises"
author: "Aaron Palumbo"
date: "November 9, 2015"
output: pdf_document
---

## 9.4.2 (Section 9.4.6)

If we wish to start out, as in Fig. 9.10, with all $U$ and $V$ entries set to the same value, what value minimizes the RMSE for the matrix $M$ of our running example?

```{r}
M = matrix(c(5, 2,  4, 4, 3, 
             3, 1,  2, 4, 1,
             2, NA, 3, 1, 4,
             2, 5,  4, 3, 5,
             4, 4,  5, 4, NA),
           byrow=TRUE, ncol=5)

table(M)
```


If $u_{ij} = v_{ij} = x$ then we have that the RMSE for $M$ is:

$$
\begin{array}{rcll}
  \text{RMSE} & = &&
  3 (1 - 2x^2)^2 + 
  4 (2 - 2x^2)^2 + 
  4 (3 - 2x^2)^2 + 
  8 (4 - 2x^2)^2 + 
  4 (5 - 2x^2)^2
  \\
  \min(\text{RMSE}) & = & \frac{d}{dx}&
  3 (1 - 2x^2)^2 + 
  4 (2 - 2x^2)^2 + 
  4 (3 - 2x^2)^2 + 
  8 (4 - 2x^2)^2 + 
  4 (5 - 2x^2)^2 = 0
  \\
  0 & = & 
  \frac{d}{dx} {\big (}&
      3  (1  - 4x^2  + 4x^4) +\\
  &&& 4  (4  - 8x^2  + 4x^4) +\\
  &&& 4  (9  - 12x^2 + 4x^4) +\\
  &&& 8  (16 - 16x^2 + 4x^4) +\\
  &&& 16 (25 - 20x^2 + 4x^4)
  {\big )}
  \\
  & = &
  \frac{d}{dx} &
  283 - 300x^2 + 92x^4
  \\
  & = && -600x + 368x^3
  \\
  & = && 8x (46x^2 - 75)
\end{array}$$

From this we have: $8x = 0 \rightarrow x = 0$  
and $46x^2 - 75 = 0 \rightarrow x = `r sqrt(75/46)`$

## 9.4.3 (Section 9.4.6)

Starting with the $U$ $V$ matrices in Fig. 9.16, do the following in order:

```{r}
returnOpt <- function(U, V, M, uij=NA, vij=NA) {
  # uij OR vij designate the element to be optimized
  if (all(is.na(uij)) + all(is.na(vij)) != 1) {
    stop("One and only one in {uij, vij} == NA")
  }
  # Are we optimizing U or V
  if (!all(is.na(uij))) {
    r <- uij[1]; s <- uij[2]
    optU <- TRUE
  } else {
    r <- vij[1]; s <- vij[2]
    optU <- FALSE
  }
  
  # return proper optimizing function
  optFun <- function(optU, r, s) {
    if (optU) {
      return(
        function(U, V, M) {
          sum(V[s, ] * (M[r, ] - U[r, ] %*% V), na.rm=TRUE) / 
            sum(V[s, ][!is.na(M[r, ])]^2)
        }
      )
    } else {
      return(
        function(U, V, M) {
          sum(U[ ,r] * (M[ ,s] - U %*% V[ ,s]), na.rm=TRUE) / 
            sum(U[ ,r][!is.na(M[ ,s])]^2)
        }
      )
    }
  }
  
  f <- optFun(optU, r, s)
  
  if (optU) {
    U[r, s] <- 0
    U[r, s] <- f(U, V, M)
    return(U)
  } else {
    V[r, s] <- 0
    V[r, s] <- f(U, V, M)
    return(V)
  }
}
```

```{r}
m <- 5
n <- 5
d <- 2

U <- matrix(rep(1, m*d), ncol=d)
V <- matrix(rep(1, d*n), nrow=d)

U <- returnOpt(U, V, M, uij=c(1, 1))
V <- returnOpt(U, V, M, vij=c(1, 1))
U <- returnOpt(U, V, M, uij=c(3, 1))
U
V
```

### (a) 9.4.3

Reconsider the value of $u_{11}$. Find it's new best value, given the changes that  have been made so far.

```{r}
U <- returnOpt(U, V, M, uij=c(1, 1))
U
```


### (b) 9.4.3

Then choose the best value for $u_{52}$.

```{r}
U <- returnOpt(U, V, M, uij=c(5, 2))
U
```


### (c) 9.4.3

Then choose the best value for $v_{22}$.

```{r}
V <- returnOpt(U, V, M, vij=c(2, 2))
V
```





